---
output: 
  github_document:
    html_preview: false
    keep_html: true
always_allow_html: true
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(webshot)
```

# vdemlite

<!-- badges: start -->
<!-- badges: end -->

`vdemlite` is a package that provides a lightweight version of the Varieties of Democracy (V-Dem) dataset that incorporates many of the core measures of V-Dem for the contemporary period and is easy to work with in R. The package provides functions for downloading and loading the data, as well as for exploring the data and searching for particular indicators.

`vdemlite` addresses two challenges that arise with the using full dataset. First, it can be readily used in cloud and production environments like Posit Cloud and shinyapps.io without having to purchase additional computing resources. Second, it provides user-friendly functions and arguments that makes it easier to download, filter, and explore the data in R. The goal is to make the dataset more accessible to new users who may not have completely developed a wrangling workflow and more convenient for more advanced users who frequently use V-Dem for contemporary analysis.

The indicators included in `vdemlite` are taken from the [Structure of V-Dem Indices, Components and Indicators](https://v-dem.net/documents/41/v-dem_structureofaggregation_v14.pdf) document that is in turn drawn from Appendix A of the V-Dem codebook.

## Installation

You can install the development version of vdemlite from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pkg_install("eteitelbaum/vdemlite")
```

## How to Use `vdemlite`

`vdemlite` has three functions: `fetchdem()`, `summarizedem()`, and `searchdem()`.

`fetchdem()` is the main function that downloads the V-Dem data and loads it into R. Within `fetchdem()`, you can specify one or more V-Dem indexes or indicators along with the years and countries you are interested in. The function will download the data and return a tibble with the specified indexes and indicators.

```{r}
library(vdemlite)

# Grab the polyarchy scores for the United States between 2000 and 2020
my_indicators <- fetchdem(indicators = "v2x_polyarchy",
                         start_year = 2000, end_year = 2020,
                         countries = "USA")
```

You can retrieve data for multiple indicators and countries by using the combine function `c()`. `fetchdem` also includes functionality for retrieving groups of indicators as they relate to the structure and aggregation of the V-Dem dataset. See the documentation for details. 

If you want a quick summary of an indicator or set of indicators, `summarizedem()` will provide an HTML table with summary statistics and missing data by country for the specified indicators.

```{r}
# summarize the civil liberties index
summarizedem("v2x_civlib")
```

Finally, you can use `searchdem` to find specific indicators or all of the indicators used to construct an index. For example, we can use `searchdem` to find all of the indicators used to construct the civil liberties index.

```{r example}
library(vdemlite)
## search for indicators that contain the word "civil liberties"
searchdem("civil liberties")
```

`searchdem()` without any arguments produces a searchable HTML table of all indicators in the dataset. 
