---
title: "Save V-Dem Lite Data"
author: "Emmanuel Teitelbaum"
format: html
---

There are some tags in the *Structure of V-Dem Indices, Components and Indicators* are missing from V-Dem. Let's check which ones are missing.

```{r}
library(dplyr)

# Use devtools:load_all() to make lookup_tbl available
devtools::load_all()

# store the tags from the lookup table in a separate object
tags <- lookup_tbl |> 
  distinct(tag) 
```

Some of the tags in the lookup table are missing in the V-Dem data. Let's see which ones are missing.

```{r}
library(vdemdata)

vdem_vars <- colnames(vdem)
missing_tags <- setdiff(tags$tag, vdem_vars)

# Display or use missing_tags as needed
print(missing_tags)
```

Now let's create a downsized version of the V-Dem dataset with only the variables that are listed in the lookup table that we derived from [Structure of V-Dem Indices, Components and Indicators](https://v-dem.net/documents/41/v-dem_structureofaggregation_v14.pdf).

```{r}
library(vdemdata)

vdemlite <- vdem |>
  select(country_name, 
         country_text_id,
         country_id,
         year, 
         # all_of() won't work because some of the tags are missing in V-Dem
         any_of(tags$tag)) |> # 
  filter(year >= 1946) 
```

Let's generate a mising variable report. 

```{r}
# Load necessary libraries
library(naniar)

# Assuming vdem_lite is your data frame
# Summarize missing data
missing_summary <- miss_var_summary(vdemlite)

# Define cutpoints for degrees of missingness
cutpoints <- c(0, 1, 5, 10, 20, 50, 100)

# Categorize variables based on the defined cutpoints
missing_summary <- missing_summary |>
  mutate(
    missing_category = cut(
      pct_miss,
      breaks = cutpoints,
      labels = c("< 1%", "1-5%", "6-10%", "11-20%", "21-50%", "51-100%"),
      include.lowest = TRUE
    )
  )

# Create a summary table
missing_table <- missing_summary |>
  group_by(missing_category) |>
  summarise(
    num_variables = n(),
    pct_variables = n()/nrow(missing_summary),
    avg_pct_missing = mean(pct_miss)
   )

# Print the summary table
missing_table
```

Let's save the V-Dem Lite dataset to a file.

Call `use_data()`:

```{r}
usethis::use_data(vdemlite, overwrite = TRUE)
file.info("../data/vdemlite.rda")$size / 1024^2
tools::resaveRdaFiles("../data/vdemlite.rda")
tools::checkRdaFiles("../data/vdemlite.rda")
file.info("../data/vdemlite.rda")$size / 1024^2
```

